Bioconductor workflow for single-cell RNA sequencing: Normalization, dimensionality reduction, clustering, and lineage inference

### 2017-12-14 workflow/workflow.Rmd
different variable genes

EP: I have only updated the code, not the text. The text needs adaptation, especially future directions and references to `SummarizedExperiment` which should be `SingleCellExperiment`.

@@

```{r packages}
# Bioconductor
library(BiocParallel)
library(clusterExperiment)
library(scone)
library(zinbwave)

# GitHub
library(slingshot)

@@

EP: There's code in SingleCellExperiment for spike-ins. We should use it, though I'm not familiar with it.

@@

EP: changing 'se' to 'dataObj' because it changes classes. Also can search and replace 'dataObj' and not 'se'.  
FP: good idea.

@@

```{r}
# # Filtering to top 1,000 most variable genes
# vars <- rowVars(log1p(assay(dataObj)))
# names(vars) <- rownames(dataObj)
# vars <- sort(vars, decreasing = TRUE)
# dataObj <- dataObj[names(vars)[1:1000],]

#alternative script:
dataObj<-makeFilterStats(dataObj,filterStats="var")
dataObj<-filterData(dataObj,percentile=1000)
```

EP: Added existing code in the clusterExperiment package that does this. Also will mean that any future steps in package that need variance for the dataObj will have variance per gene stored.

@@

@@

EP: This saves a lot of trouble with the plotHeatmap, etc.

@@

FP: W can now be accessed using reducedDim function. Two functions called reducedDim are loaded in the workflow: one from slingshot and the other one from singleCellExperiment.

EP: I would get rid of the `SingleCellExperiment::` code and just add a `library(SingleCellExperiment)`. Much more comfortable code for most users.

```{r mdsW, fig.cap="ZINB-WaVE: MDS of the low-dimensional matrix W, where each point represents a cell and cells are color-coded by original published clustering."}
W <- SingleCellExperiment::reducedDim(dataObj)
d <- dist(W)
fit <- cmdscale(d, eig = TRUE, k = 2)
plot(fit$points, col = col_clus[as.character(publishedClusters)], main = "",
     pch = 20, xlab = "Component 1", ylab = "Component 2")
legend(x = "topleft", legend = unique(names(col_clus)), cex = .5, fill = unique(col_clus), title = "Sample")
```

@@

```{r addW}
#SingleCellExperiment::reducedDim(dataObj,"zinbwave50")<-W
```

@@

@@

```{r boxplotReg3g, fig.width=6, fig.height=4, fig.cap="RSEC: Boxplots of the log count of gene Reg3g stratified by cluster."}
plotFeatureBoxplot(dataObj, whichCluster="primary", feature="Reg3g",ylab = "Reg3g count values", cex.axis = .8, cex.lab = .8,assay=log1p(assays(dataObj)[["counts"]]))
# plotFeatureBoxplot(dataObj, whichCluster="primary", feature="Reg3g",ylab = "Reg3g normalized Expression values", cex.axis = .8, cex.lab = .8)

```

@@

```{r slingshotsupervised, eval=FALSE}
lineages <- getLineages(X, clusterLabels = primaryClusterNamed(pseudoCe), start.clus = "c1",
                         end.clus = c("c3", "c7"))
lineagees <- getCurves(lineages)
pairs(lineages, type="curves", col = colorCl)
pairs(lineages, type="lineages", col = colorCl, 
      show.constraints = TRUE)

>>>>

EP: I have only updated the code, not the text. The text needs adaptation, especially future directions and references to `SummarizedExperiment` which should be `SingleCellExperiment`.
FP: TODO, fix ref.

@@

```{r packages}
# GitHub
library(slingshot)

# Bioconductor
library(BiocParallel)
library(clusterExperiment)
library(scone)
library(zinbwave)
library(SingleCellExperiment)

@@

EP: There's code in SingleCellExperiment for spike-ins. We should use it, though I'm not familiar with it. 
FP: Davide, what is the standard way to use SingleCellExperiment for spike-ins?

@@

@@

```{r}
# # Filtering to top 1,000 most variable genes
vars <- rowVars(log1p(assay(dataObj)))
names(vars) <- rownames(dataObj)
vars <- sort(vars, decreasing = TRUE)
# dataObj <- dataObj[names(vars)[1:1000],]

#alternative script:
dataObj<-makeFilterStats(dataObj,filterStats="var")
dataObj<-filterData(dataObj,percentile=1000,filterStats="var")

length(intersect(rownames(dataObj), names(vars)[1:1000]))
setdiff(rownames(dataObj), names(vars)[1:1000])
```

EP: Added existing code in the clusterExperiment package that does this. Also will mean that any future steps in package that need variance for the dataObj will have variance per gene stored.  
FP: great. I added the filterStats argument to filterData function. I get an error otherwise.

FP: We get a different list of most variable genes here.

@@

FP: These values are slightly different from what

@@

EP: This saves a lot of trouble with the plotHeatmap, etc.
FP: Ok.

@@

```{r mdsW, fig.cap="ZINB-WaVE: MDS of the low-dimensional matrix W, where each point represents a cell and cells are color-coded by original published clustering."}
W <- reducedDim(dataObj)
d <- dist(W)
fit <- cmdscale(d, eig = TRUE, k = 2)
plot(fit$points, col = col_clus[as.character(publishedClusters)], main = "",
     pch = 20, xlab = "Component 1", ylab = "Component 2")
legend(x = "topleft", legend = unique(names(col_clus)), cex = .5, fill = unique(col_clus), title = "Sample")
```

@@

```{r addW}
#reducedDim(dataObj,"zinbwave50")<-W
```

@@

FP: We get more clusters for the combineMany method than before. Maybe because the most variable genes are different?

@@

```{r boxplotReg3g, fig.width=6, fig.height=4, fig.cap="RSEC: Boxplots of the log count of gene Reg3g stratified by cluster."}
plotFeatureBoxplot(dataObj, whichCluster="primary", feature="Reg3g",ylab = "Reg3g count values", cex.axis = .8, cex.lab = .8,assay=log1p(assays(dataObj)[["counts"]]))
```

@@

```{r slingshotsupervised, eval=FALSE}
lineages <- getLineages(X, clusterLabels = primaryClusterNamed(pseudoCe), start.clus = "c1",
                         end.clus = c("c3", "c7"))
lineages <- getCurves(lineages)
pairs(lineages, type="curves", col = colorCl)
pairs(lineages, type="lineages", col = colorCl, 
      show.constraints = TRUE)
